---
title: "Data Analysis and Visualisation with R"
format: html
---

## Learning objectives

* Get familiar with an IDE.
* Know how to set up a data research project.
* Be able to read in, process, and store data.
* Visualise data: choose an appropriate visualisation for your data.


## Data Visualization Notes
For this workshop, we borrow heavily from [Data Visualization for Social Science](https://socviz.co/index.html#preface).
This very markdown is based off the R package accompanying the book. 
You can use it to take notes, write your code, and produce a good-looking, reproducible document that records the work you have done. 

At the very top of the file is a section of *metadata*, or information about what the file is and what it does. 
The metadata is delimited by three dashes at the start and another three at the end.
You should change the title, author, and date to the values that suit you. 
Keep the `output` line as it is for now, however. 
Each line in the metadata has a structure. 
First the *key* ("title", "author", etc), then a colon, and then the *value* associated with the key.  
It is very picky when it comes with indentations, characters used, etc.


## This is a Quarto File
[Markdown](http://rmarkdown.rstudio.com) is a simple formatting syntax for authoring HTML, PDF, and MS Word documents, and is also the basis for [Quarto](https://quarto.org/). 

When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. 
A *code chunk* is a specially delimited section of the file. 
You can add one by moving the cursor to a blank line choosing Code > Insert Chunk from the RStudio menu. 
When you do, an empty chunk will appear:
```{r}
```

Code chunks are delimited by three backticks (found to the left of the 1 key on US and UK keyboards) at the start and end.
The opening backticks also have a pair of braces and the letter `r`, to indicate what language the chunk is written in. 
You write your code inside the code chunks. 
Write your notes and other material around them, as here. 


## Know your editor
Please have a look at your screen.
Some of the most important things to learn now:

* Console
* Editor
* Environment


# Before you begin to explore and analyse, set up your workspace
A number of functions and packages come already install in what is generally referred to as `R base` (and you can actually see what's in it by typing `base::` in your console, and let the auto-complete suggest you the full list of functions ... ).
However, most of the things we'll do in this workshop need further `libraries`.
To install the `tidyverse`, make sure you have an Internet connection. 
Then *manually* run the code in the chunk below. 
If you knit the document if will be skipped. 
We do this because you only need to install these packages once, not every time you run this file. 
Either knit the chunk using the little green "play" arrow to the right of the chunk area, or copy and paste the text into the console window.
```{r install, eval = FALSE}

## This code will not be evaluated automatically. (Notice the eval = FALSE declaration in the options section of the code chunk)

my_packages <- c("tidyverse", "broom", "coefplot", "cowplot",
                 "gapminder", "GGally", "ggrepel", "ggridges", "gridExtra",
                 "here", "interplot", "margins", "maps", "mapproj",
                 "mapdata", "MASS", "quantreg", "rlang", "scales",
                 "survey", "srvyr", "viridis", "viridisLite", "devtools")

install.packages(my_packages, repos = "http://cran.rstudio.com")
devtools::install_github("kjhealy/socviz")
```


## Set Up Your Project and Load Libraries
To begin we must load some libraries we will be using. 
If we do not load them, R will not be able to find the functions contained in these libraries. 
The tidyverse includes ggplot and other tools. 
We also load the socviz and gapminder libraries.
```{r setup}
#| include: false

## Load the libraries we will be using
library(tidyverse)
library(here)
library(gapminder)
library(ggrepel)
library(socviz)
```

Notice that here, the braces at the start of the code chunk have some additional options set in them. 
There is the language, `r`, as before. 
This is required. 
Then there is the word `setup`, which is a label for your code chunk.
Labels are useful to briefly say what the chunk does. 
Label names must be unique (no two chunks in the same document can have the same label) and cannot contain spaces. 
Then, after the comma, an option is set: `include=FALSE`. 
This tells R to run this code but not to include the output in the final document. 

When you click the **Render** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
You can embed an R code chunk like this:

```{r}
gapminder
```


## Get Started
Two ways to create a plot.

* We can assign a ggplot plot to an object called `p`, and then add (literally, using `+`) layers on top of it.
```{r}
p <- ggplot(data= gapminder,
          mapping = aes(x= gdpPercap,
                        y= lifeExp))
p + geom_point()
```

* Another way of achieving the same thing is as follows, which is more in line with a tidyverse style which heavily rely on pipes (`|>`). 
```{r}
gapminder |>
  ggplot(aes(x= gdpPercap, y= lifeExp)) + 
  geom_point()
```

Integrating from 'R for Data Science'
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ,
                           y = hwy))
```

geom_smooth() using method = 'gam' and formula 'y ~ s(x, bs = "cs")'; GAM stands for generalised additive model
```{r}
p + geom_smooth()
p + geom_point() + geom_smooth() 
```

Using method='lm' (linear model) as an argument to geom_smooth()
```{r}
p<-ggplot(data= gapminder,
          mapping = aes(x= gdpPercap,
                        y= lifeExp))
p + geom_point() + geom_smooth(method = 'lm')
```

Data is quite bunched up against the left side. Gross Domestic Product per capita is not normally distributed across our country years. The x-axis scale would probably look better if it were transformed from a linear scale to a log scale. For this we can use a function called scale_x_log10(). As you might expect this function scales the x-axis of a plot to a log 10 basis. To use it we just add it to the plot:
```{r}
p<-ggplot(data= gapminder,
          mapping = aes(x= gdpPercap,
                        y= lifeExp))
p + geom_point() + 
  geom_smooth(method = 'gam') +
  scale_x_log10()
```

The labels on the tick-marks can be controlled through the scale_ functions.
```{r}
p + geom_point() + 
  geom_smooth(method = 'gam') +
  scale_x_log10(labels = scales:: dollar)
```

See the difference between this (WRONG)...
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = "purple"))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

And this (CORRECT)
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_point(color = "purple") +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

For a variation, see
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) 
p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = FALSE, size = 1, method = "lm") +
    scale_x_log10()
```

For a second variation, see
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) 
p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = FALSE, size = 1, method = "lm") +
    scale_x_log10()
```

The SE option commands the standard error, and we can switch it off by inputing se= FALSE. THE ALPHA property regualte the transparency of the objects, from 0 to 1. for a more colourful version of th graph above, with continent colours differently, see 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color= continent)) 
p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = FALSE, size = 1, method = "lm") +
    scale_x_log10()
```

To add titles and captions, use labs as following

```{r}
p<-ggplot(data=gapminder,
          mapping= aes(x= gdpPercap,
                       y= lifeExp))
p + geom_point(alpha= 0.3) +
  geom_smooth(method = 'gam') +
  scale_x_log10(labels= scales:: dollar) +
  labs(x= 'GDP per capita', 
       y= 'Life Expectancy in Years', 
       title= 'Economic Growth and Life Expectancy', 
       subtitle= 'Data points are country-years', 
       caption= 'Source: Gapminder.')
```

For a variation, see
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color= continent)) 
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

to fill the ribbon of the standard error, usee fill
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color= continent,
                          fill= continent)) 
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

To map only the continents, but not the lines or the ribbon, and cutting down on the lines will make the graph more readable
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) 
p + geom_point(mapping= aes(color= continent)) +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

To map continuous variables in gradient, use the color function not in the ggplot section, but in the geom one, exactly as follows
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) 
p + geom_point(mapping= aes(color= log(pop))) +
    scale_x_log10()
```


unvotes

ep data (documents)

polls (EE, de, means, aggregations)


